---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bifurcationEWS
bifurcationEWS is an R package for running an Early Warning Signal analysis to anticipate deterministic bifurcations in dynamical systems. The package supports timeseries simulation of many deterministic bifurcations, computation of some univariate and multivariate EWS, and an assessment of the performance of EWS. 

<!-- badges: start -->
<!-- badges: end -->

## Installation

You can install the development version of bifurcationEWS from [GitHub](https://github.com/) with:

```{r}
# install.packages("devtools")
devtools::install_github("KCEvers/bifurcationEWS")
library(bifurcationEWS)

```

## Generate timeseries

Let's generate some timeseries for a deterministic Hopf bifurcation:

```{r }
# Define control parameter
bifpar_pars = list(
        bifpar_start = .75,
        bifpar_end = .85,
        pre_steps = 0,
        baseline_steps = 0,
        transition_steps = 51,
        post_steps = 0
      )
# Generate Hopf bifurcation timeseries
GLV_Hopf = bifurcation_ts(model = GLV_model,
                         bifpar_pars = bifpar_pars,
                         nr_timesteps = 1000
                        )
df = GLV_Hopf$df

```

```{r }

# regimes_Hopf = find_regimes(GLV_Hopf_trans, factor_k = 1)
# plot_bifdiag(regimes_Hopf, sel_variables = "X1")

```

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. -->

<!-- You can also embed plots, for example: -->

<!-- ```{r pressure, echo = FALSE} -->
<!-- plot(pressure) -->
<!-- ``` -->

<!-- In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN. -->
