---
title: "Demo of bifurcationEWS package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Demo of bifurcationEWS package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, echo=FALSE,
  # fig.width=10, fig.height=2, 
  fig.fullwidth=TRUE,
  comment = "#>"
)
rgl::setupKnitr(autoprint = TRUE) # suppress opening new window
```

```{r setup}
# devtools::install_github("KCEvers/bifurcationEWS")
library(bifurcationEWS)
library(ggh4x)
library(dplyr)
library(plotly)
library(ggplot2)
library(foreach)
```

### Generating timeseries

```{r demo GLV}
pars <- setup_pars(model_name = "detGLV")
timestep = .01
nr_timesteps = 1000
GLVs = list()
ss =  c(.85, .94, .969, .99, 1.08, .175)
for (s_idx in seq_along(ss)){
  s = ss[s_idx]
  bifpar_list <- list(list("s" = s))
  GLV = bifurcation_ts(model = GLV_model, model_pars = pars$model_pars,
                 bifpar_list = bifpar_list,
                 timestep = timestep, nr_timesteps = nr_timesteps,
                 # downsample_pars = downsample_pars,
                 X_names = pars$X_names,
                 # X0 = c(X1=0,X2=runif(1),X3=runif(1), X4=runif(1))
                 seed_nr = round(s*100)
                 )
  GLVs[[s_idx]] = GLV$df %>% dplyr::mutate(s = sprintf("%.4f", s), bifpar_idx = s_idx)
}

GLV_dfs = GLVs %>% do.call(rbind, .) %>% as.data.frame()
head(GLV_dfs)

# Timeseries plot
pl_ts = GLV_dfs %>% 
  ggplot() + 
  geom_line(aes(x = time_idx, y = X1, color = factor(s)), linewidth = .1) +
  geom_point(aes(x = time_idx, y = X1, color = factor(s)), size = .1) +
    labs(x = 'Time', y = "X1", title = "Timeseries") + 
  facet_wrap(s ~ ., ncol = 1) +
  viridis::scale_color_viridis(discrete=TRUE, name = "Bifurcation parameter")
  
print(style_plot(pl_ts))


landscape_GLVs = plot_3D_landscape_trans(GLV_dfs,
                                         bifpar_idxs = seq_along(ss),
                                         bifpar_values = ss
                                         # x1_range = c(0,.5), x2_range = c(0.25,0.75),
                                         # x3_range = c(0,0.5)
                                         )
print(landscape_GLVs)
```


What are some more complicated regimes that we can find in the model?


### Bifurcation diagram

```{r bifurcation diagram, include=FALSE}
# # Generate data with changing bifurcation parameter
# downsample_pars = list(type = c("average", "one_sample")[1],
#                                                   win_size = 50,
#                                                   which_X = c(50, "first", "middle", "last", "random")[1],
#                                                   seed = 123)
timestep = .01
nr_timesteps = 1000
bifpar_pars = list(
#   bifpar_start = 1.4, bifpar_end = .6,
  bifpar_start = .96, bifpar_end = .97,
#   #bifpar_start = .965, bifpar_end = .985,
                                    pre_steps = 0,
                                  baseline_steps = 0, transition_steps = 50, post_steps = 0)
bifpar_list=do.call(get_bifurcation_range, bifpar_pars)

GLV_Chaos1 = bifurcation_ts(model = GLV_model, model_pars = pars$model_pars,
               bifpar_list = bifpar_list,
               timestep = timestep, nr_timesteps = nr_timesteps,
#                downsample_pars = downsample_pars,
               X_names = pars$X_names,
               seed_nr = 100
               )
# GLV_full = bifurcationEWS::GLV_full

```


```{r plots, echo=FALSE}
pl_ts = GLV_Chaos1$df %>% 
  dplyr::filter(bifpar_idx %in% c(2)) %>%
  ggplot() + 
  geom_line(aes(x = time_idx, y = X1), linewidth = .1) +
  geom_point(aes(x = time_idx, y = X1), size = .1) +
    labs(x = 'Time', y = "X1", title = "Timeseries") 
print(style_plot(pl_ts))

bifpar_idxs = ceiling(round(length(bifpar_list) * seq(.2, 1, by = .2))) %>% unique
landscape_GLV_Chaos1 = plot_3D_landscape_trans(GLV_Chaos1$df,
                                         bifpar_idxs = bifpar_idxs,
                                         bifpar_values = unlist(purrr::map(bifpar_list[bifpar_idxs], "s"))

                                         # x1_range = c(0,.5), x2_range = c(0.25,0.75),
                                         # x3_range = c(0,0.5)
                                         )
print(landscape_GLV_Chaos1)

# Get dataframe with peaks
peaks_df = peaks_bifdiag(GLV_Chaos1$df, GLV_Chaos1$X_names)
# head(peaks_df)

# peaks_df_ = peaks_df %>% dplyr::filter(variable == "X1", bifpar_idx > 1)
pl_peaks = peaks_df %>% ggplot() + geom_point(aes(x = bifpar_idx, y = X), size = .1) +
  labs(x = 'Bifurcation parameter', y = "X", title = "Transient bifurcation diagram") +
  ggh4x::facet_grid2(variable ~ .)

print(style_plot(pl_peaks))
```

### Finding regime boundaries

How do we find the boundaries between these two regimes? A first approach might be to simply count the number of distinct peaks for each bifurcation parameter. However, this neglects the temporal sequence of these peaks, which is required to be consistenct for a limit cycle. Change point detection algorithms would be another option, yet these are often not suitable for cyclical data.

The code below executes a quite crude yet effective approach. For a number of candidate period lengths $k \in ks$, we cluster the peaks corresponding to the same time in the period. We then compute the within-cluster sum of squares (WCSS) to get a sense of the spread within each cluster. A period length $k$ with minimum WCSS is determined as the best period (we deal with subharmonics in a subsequent step, see code).

Note that although there are much more sophisticated temporal clustering algorithms out there (such as provided in the R package **dtwclust**), we only required a very simple metric to assess how well temporal sequences could be distinguished.

```{r periodicity, echo=TRUE}
regimes_full = find_regimes(GLV_Chaos1,
                             thresh_node = .001,
                             thresh_coord_spread = .025,
                             thresh_peak_idx_spread=2,
                             min_length_regime = 1,
                             nr_smooth = 0,
                              factor_k = .1, 
                            remove_initial_bifpar_idx = 1
)

print(regimes_full$regimes)
print(regimes_full$regime_bounds)
```

```{r echo=FALSE}
pl_regimes = regimes_full$peaks_df %>% ggplot() +
    geom_rect(data=regimes_full$regimes,aes(xmin=start_bifpar_idx,xmax=end_bifpar_idx,
                                           ymin=-Inf, ymax=Inf, fill = regime), linewidth = .5, alpha = .4) +
    geom_vline(data = regimes_full$regime_bounds, aes(
      xintercept = regime1_end_idx,
      col = regime_switch_type), linewidth = 1.3, linetype = 'dashed') +
    geom_vline(data = regimes_full$regime_bounds, aes(
      xintercept = regime2_start_idx,
      col = regime_switch_type), linewidth = 1.3, linetype = 'dashed') +
    geom_point(aes(x = bifpar_idx, y = X),col='grey30',alpha = .8, size = .25) +
    viridis::scale_color_viridis(
      discrete=TRUE,
      option = 'magma',
      begin = 0, end = .8, name = "Regime switch (end of first and start of second regime)") +
    viridis::scale_fill_viridis(
      discrete=TRUE,
      option = 'turbo',
      begin = 0, end = 1, name = "Period",
      guide = guide_legend(override.aes = list(alpha = 1,
                                               size = 3))) +
    guides(color = guide_legend(ncol = 1), fill = guide_legend(ncol = 1)) + # x rows in legend
    ggh4x::facet_grid2(variable ~ .) +
    labs(x = 'Bifurcation parameter', y = "", title = "Transient bifurcation diagram")

print(style_plot(pl_regimes))
```

### Generating null and transition models

Now, we might be interested in a select few transitions. To get a sense of the sensitivity and specificity of our ability to detect these transitions, we need both transition and null models. To fairly compare transitions, we make them of the same time length, yet this does also go hand in hand with a change in the rate of the control parameter. 

Say we are interested in the regime switch of *** - how might we simulate this? We already have the range of $s$ for which a particular regime switch occurs. However, the system is also multistable for many parameter ranges, such that a randomly chosen initial condition is not likely to yield the desired regime. To circumvent this issue, we make use of our full GLV timeseries. This timeseries offers both the regime boundaries $s$ and corresponding initial condition needed to simulate a new (much shorter) transition model. In addition, we can now add a proper baseline and transition period. 



```{r echo=FALSE}
# Select regime switch
GLV_full = bifurcationEWS::GLV_full
regimes_full = find_regimes(GLV_full,
                             thresh_node = .001,
                             thresh_coord_spread = .025,
                             thresh_peak_idx_spread=2,
                             min_length_regime =5,
                             nr_smooth = 0,
                              factor_k = .1)

regime_switch = "PD_4to8"
desired_regime1 = "Period-4 (X2,X3,X4) AND Period-6 (X1)" 
desired_regime2 = "Period-10 (X1) AND Period-8 (X2,X3,X4)"
desired_switch = regimes_full$regime_bounds %>%
  # dplyr::filter(regime_switch_type == "Mixed-Periodic to Chaotic or Transitioning" & regime1 == "Period-16 (X2,X3,X4) AND Period-20 (X1)" & regime2 == "Chaotic or Transitioning")
  dplyr::filter(
    regime1 == !!desired_regime1 &
      regime2 == !!desired_regime2
  )

bifpar_pars_PD_4to8 = list(
  pre_steps = 50,
  baseline_steps = 50,
  transition_steps = 50,
  post_steps = 50,
  bifpar_start = GLV_full$bifpar_list[[desired_switch$regime1_halfway_idx]]$s,
  bifpar_end = GLV_full$bifpar_list[[desired_switch$regime2_start_idx]]$s
)

GLVs = list()
regime_lists = list()


for (trans_or_null in c("transition", "null")) {
    if (trans_or_null == "transition") {
      bifpar_pars_PD_4to8_ = bifpar_pars_PD_4to8
    } else if (trans_or_null == "null") {
      bifpar_pars_PD_4to8_ = utils::modifyList(bifpar_pars_PD_4to8, list(bifpar_end = NA))
    }
    X0 = unlist(desired_switch[, GLV_full$X_names])
    
    # Generate data with changing bifurcation parameter
    # GLV = bifurcation_ts(model = GLV_model, 
    # model_pars = GLV_full$model_pars,
    #                bifpar_pars = bifpar_pars_PD_4to8,
    #                X0 = X0,
    #                timestep = GLV_full$timestep,
    #                 nr_timesteps = GLV_full$nr_timesteps,
    #                 downsample_pars = GLV_full$downsample_pars,
    #                X_names = GLV_full$X_names
    #                )
    # Use package-made GLV
    if (trans_or_null == "transition") {
      GLV = bifurcationEWS::GLV_trans
    } else if (trans_or_null == "null") {
      GLV = bifurcationEWS::GLV_null
    }
    head(GLV$df)

    # We need to check whether our simulation generated the desired regimes using the same process as above:
   
     # Get dataframe with peaks
    peaks_df = peaks_bifdiag(GLV$df, GLV$X_names)
    regime_list = find_regimes(
      GLV,
      thresh_node = .001,
      thresh_coord_spread = .025,
      thresh_peak_idx_spread = 2,
      min_length_regime = 5,
      nr_smooth = 0,
      factor_k = .1
    )
    # regime_list = 
    # peaks_df = regime_list$peaks_df
    print(regime_list$regimes)
    print(regime_list$regime_bounds)
        
#     pl_ts_trans = GLV$df %>%
#       dplyr::filter(bifpar_idx == 2) %>%
#       tidyr::gather(variable, value,-setdiff(colnames(.), GLV$X_names)) %>%
#       ggplot() + geom_point(aes(x = time, y = value), size = .1) +
#       ggh4x::facet_grid2(variable ~ .)
#     print(style_plot(pl_ts_trans))
#     
#     pl_peaks = peaks_df %>% dplyr::filter(variable == "X1") %>%
#       ggplot() + geom_point(aes(x = bifpar_idx, y = X), size = .1) +
#       labs(x = 'Bifurcation parameter', y = latex2exp::TeX("$X_1$"), 
#            title = "Transient bifurcation diagram") +
#       geom_rect(aes(xmin = pre_steps + baseline_steps, 
#                     xmax = pre_steps + baseline_steps + transition_steps,
#                     ymin = -Inf, ymax = Inf
#     ),
#     alpha = 0.85, col = "blue", fill = NA,
#     linewidth = 1.25) +
#             geom_rect(data = regime_list$regime_bounds %>% dplyr::filter(regime1 == "Period-4 (X2,X3,X4) AND Period-6 (X1)"), 
#                 aes(xmin = regime2_start_idx - 1 - transition_steps, 
#                     xmax = regime2_start_idx - 1,
#                     ymin = -Inf, ymax = Inf
#     ),
#     alpha = 0.85, col = "green", fill=NA,
#     linewidth = 1.25) 
#       # geom_vline(data = data.frame(
#       #   xintercept = c(
#       #     pre_steps + baseline_steps,
#       #     pre_steps + baseline_steps + transition_steps
#       #   ),
#       #   moment = c("Baseline", "Transition"),
#       #   col = rep("Parameter change",2)
#       # ),
#       # aes(xintercept = xintercept))
#     
#     print(style_plot(pl_peaks))
#      
#     # Plot change in phase space
# bifpar_idxs = c(pre_steps + baseline_steps, pre_steps + baseline_steps + round(.75 * transition_steps),
#                 pre_steps + baseline_steps + transition_steps + 1)
#    landscape_GLV = plot_3D_landscape_trans(GLV$df,
#                                          bifpar_idxs = bifpar_idxs,
#                                          bifpar_values = unlist(purrr::map(GLV$bifpar_list[bifpar_idxs], "s"))
#                                          # x1_range = c(0,.5), x2_range = c(0.25,0.75),
#                                          # x3_range = c(0,0.5)
#                                          )
# print(landscape_GLV)
   # Store
    GLVs[[trans_or_null]] = GLV
    regime_list[["regime_bounds"]] = regime_list[["regime_bounds"]] %>% add_par_as_cols(bifpar_pars_PD_4to8)
    regime_lists[[trans_or_null]] = regime_list
  }


regime_bounds_trans = regime_lists[["transition"]]$regime_bounds %>%
  dplyr::filter(
    regime1 == !!desired_regime1 &
      regime2 == !!desired_regime2
  ) %>%
  dplyr::mutate_at(c("transition_steps", "baseline_steps"), ~as.numeric(as.character(.))) %>%
  dplyr::mutate(transition_end_idx = na.omit(unique(regime2_start_idx)) - 1) %>%
  ungroup() %>%
  dplyr::mutate(
    transition_start_idx = transition_end_idx - transition_steps + 1,
    baseline_end_idx = transition_start_idx - 1,
    baseline_start_idx = baseline_end_idx - baseline_steps + 1,
    par_change_start_idx = ifelse(
      trans_or_null == "transition",
      pars_template$pre_steps + pars_template$default_baseline_steps,
      NA
    ),
    par_change_end_idx = ifelse(
      trans_or_null == "transition",
      pars_template$pre_steps + pars_template$default_baseline_steps + transition_steps,
      NA
    )
  )

```

The bifurcation diagram above illustrates why it is so difficult to generate regime switches: the exact timing of the regime switch may vary depending on the initial condition and rate of change in the bifurcation parameter. We use the "empirical" (and not the "intended") transition moment for the detection of EWS.

To make the regime switch harder to detect, we add some simple white observational noise.

```{r Adding observational noise}
sigma_obs_noise = .01
GLV_trans$noisy_df = GLV_trans$df %>%
  add_obs_noise(GLV_trans$X_names, noise_mean = 0, noise_sigma = sigma_obs_noise, noise_constant = 0)
pl_noisy = ggplot(data = GLV_trans$df %>% 
  dplyr::filter(bifpar_idx > 5, bifpar_idx < 7) %>%
  tidyr::gather(variable, value, -setdiff(colnames(.), GLV_trans$X_names))) +
  geom_point(aes(x = time_idx, y = value), col = 'grey50', alpha = .75, size = .05) +
  geom_point(data = GLV_trans$noisy_df %>% 
                 dplyr::filter(bifpar_idx > 5, bifpar_idx < 7) %>%
               tidyr::gather(variable, value, -setdiff(colnames(.), GLV_trans$X_names)), 
  aes(x = time_idx, y = value), col = 'orange3', alpha = .75, size = .05) +
  ggh4x::facet_grid2(variable ~ .) + labs(title = sprintf("sigma = %.2f", sigma_obs_noise),
                                          x = 'Time', y = "X1") 
print(style_plot(pl_noisy))
``` 



## Early Warning Signals

### Computing windowed EWS
```{r Windowed EWS}
uni_metrics = c(
  # "mean" = mean,
  # "variance" = var,
  # "autocorrelation" = get_autocorr,
  # "skewness" = moments::skewness,
  # "kurtosis" = moments::kurtosis,
  # "Smax" = get_Smax
)
multi_metrics = c(
  # "meanAbsCrossCorr" = get_conn,
  # "largestLambdaCovMatrix" = eigenvalue,
  # "spatial_variance" = spatial_variance,
  # "spatial_skewness" = spatial_skewness,
  # "spatial_kurtosis" = spatial_kurtosis
  "RQA" = runRQA
)
EWS_args=list()
EWS_args=list("Smax" = list(
    fs = 1 / GLV_trans$timestep / GLV_trans$downsample_pars$win_size,
    nr_timesteps = GLV_trans$nr_timesteps
  ))
split_df_EWS = run_bifEWS(GLV_trans$noisy_df, GLV_trans$X_names, uni_metrics, multi_metrics, EWS_args = EWS_args)
pl_EWS <- split_df_EWS %>% ggplot() +
    geom_line(aes(x = bifpar_idx, y = value), linewidth = .45, alpha = .5, color ='grey50') +
          geom_point(aes(x = bifpar_idx, y = value), size = .6, color ='grey20') +
  ggh4x::facet_wrap2(. ~ metric, ncol = 2, scales = "free_y",
                     # labeller = labeller(metric = label_func)
                     )
print(style_plot(pl_EWS))

split_df_EWS %>% dplyr::filter(metric == "Smax_Smax_var1")
```


### Warnings

At what point do we decide the system is transitioning?

We make use of a simple classical technique that involves **. 
By varying the width of the confidence bands, we get a better understanding of the sensitivity and specificity of the measure. 
The only two settings that are required are the definition of the baseline period and the range of $\sigma_{crit}$ that is tested.

```{r warning}

# Add confidence band
sigma_crit = 2
sigmas_crit = seq(.025, 6, by = 0.025)

# Count baseline back
baseline_idx = round(seq(regime_bounds_trans$baseline_start_idx, regime_bounds_trans$baseline_end_idx))
transition_idx = round(seq(regime_bounds_trans$transition_start_idx, regime_bounds_trans$transition_end_idx))

# Compute at which critical sigma value a warning occurs
split_df_EWS_CI = get_warnings(split_df_EWS,
                            baseline_idx,
                            transition_idx,
                                sigmas_crit = sigmas_crit,
                                nr_consecutive_warnings = 1)
  
warning_df = split_df_EWS_CI$warning_df 

# Plot
sigma_crit = 2
sz_point_nowarning = .1
sz_point_warning = 1

pl_warnings <- split_df_EWS_CI$winEWS_df %>%
      dplyr::mutate(ymin = mean_w0 - sigma_crit * sd_w0,
                    ymax = mean_w0 + sigma_crit * sd_w0,
                    value_within_bounds = ifelse(value >= ymin & value <= ymax, value, NA),
                    value_outside_bounds_on_time = ifelse((value < ymin | value > ymax) & bifpar_idx > bifpar_pars_PD_4to8$baseline_steps & bifpar_idx <= (bifpar_pars_PD_4to8$baseline_steps + bifpar_pars_PD_4to8$transition_steps), value, NA),
                    value_outside_bounds_baseline = ifelse((value < ymin | value > ymax) & bifpar_idx <= bifpar_pars_PD_4to8$baseline_steps, value, NA),
                    value_outside_bounds_too_late = ifelse((value < ymin | value > ymax) & bifpar_idx >= (bifpar_pars_PD_4to8$baseline_steps + bifpar_pars_PD_4to8$transition_steps), value, NA)
      ) %>%
      ggplot()  +
      geom_ribbon(aes(x = bifpar_idx, ymin = ymin, ymax =ymax), fill="gray80", color = "gray40", alpha = .65) +
     geom_line(aes(x = bifpar_idx, y =  as.numeric(as.character(value))), colour = "gray50", linewidth = .1) +
      geom_point(aes(x = bifpar_idx, y =  as.numeric(as.character(value_within_bounds))), colour = "gray50", size = sz_point_nowarning, alpha = .75) +
      geom_point(aes(x = bifpar_idx, y =  as.numeric(as.character(value_outside_bounds_on_time))), colour = 'red', size = sz_point_warning) +
      geom_point(aes(x = bifpar_idx, y = as.numeric(as.character(value_outside_bounds_baseline))), colour = 'purple', size = sz_point_warning) +
      geom_point(aes(x = bifpar_idx, y = as.numeric(as.character(value_outside_bounds_too_late))), colour = 'blue', size = sz_point_warning) +
      ggh4x::facet_wrap2(metric ~ .,
                         # dir = "v", strip.position = "right",
                         labeller = label_parsed,
                         # strip = ggh4x::strip_nested(bleed = TRUE),
                         ncol = 4,
                         scales = "free_y")
print(style_plot(pl_warnings))

    
```

### Sensitivity vs. Specificity

We repeat the procedure for different initialisations of added white noise to get a false and true positive **rate**.

```{r AUC}
forloop_ = get_forloop(noise_iter=1:10, 
            trans_or_null = c("transition", "null"),
            sigma_obs_noise = c(0,.02, .04), 
            regime_switch = regime_switch) %>% 
  purrr::map(function(x){
  if (x$sigma_obs_noise == 0 & x$noise_iter != 1) {
    return(NULL)
  } else {
    return(x)
  }
})
forloop = forloop_[!unlist(purrr::map(forloop_, is.null))]

EWS_warnings <- foreach(for_par = forloop, .combine = 'rbind') %do% {
    trans_or_null = for_par$trans_or_null
    noise_iter = for_par$noise_iter
    sigma_obs_noise = for_par$sigma_obs_noise
    print(as.data.frame(for_par))
    
      GLV = GLVs[[trans_or_null]]
      df = GLVs[[trans_or_null]]$df
      regime_list = regime_lists[[trans_or_null]]    
  
      # Noisy data
      noisy_df = df %>%
      add_obs_noise(GLV$X_names, noise_mean = 0, noise_sigma = sigma_obs_noise, noise_constant = 0, seed_nr = noise_iter)
    
      # EWS
      split_df_EWS = run_bifEWS(noisy_df, GLV_trans$X_names, uni_metrics, multi_metrics, EWS_args = list())
  
      # Warnings
      baseline_idx = round(seq(regime_bounds_trans$baseline_start_idx, regime_bounds_trans$baseline_end_idx)) # Count baseline back
      transition_idx = round(seq(regime_bounds_trans$transition_start_idx, regime_bounds_trans$transition_end_idx))
  
      # Compute at which critical sigma value a warning occurs
      split_df_EWS_CI = get_warnings(split_df_EWS,
                                baseline_idx,
                                transition_idx,
                                    sigmas_crit = sigmas_crit,
                                    nr_consecutive_warnings = 1)
      
     return(warning_df %>% dplyr::bind_cols(as.data.frame(for_par)))
  }

head(EWS_warnings)

# Compute ROC & AUC
grouping_vars = setdiff(names(forloop[[1]]), c("noise_iter", "trans_or_null"))
EWS_warnings_ROC = warnings_to_ROC(EWS_warnings, grouping_vars)
EWS_warnings_AUC = ROC_to_AUC(EWS_warnings_ROC, grouping_vars)

```

```{r}

plot_ROC <- function(EWS_warnings_ROC_sub, grouping_keys){
  filename = grouping_keys %>% t() %>% as.data.frame()%>%
    tibble::rownames_to_column() %>%
    dplyr::mutate_all (~ trimws(.x))%>%
    apply(1, paste0, collapse = "_") %>% paste0(collapse="_")

  pl_ROC = EWS_warnings_ROC_sub %>%
    select(-c("sum_tp", "sum_fp", "sum_tn", "sum_fn", "acc")) %>%
    dplyr::mutate_at(c("fnr", "fpr", "tnr", "tpr"), ~ as.numeric(as.character(.))) %>%
    tidyr::gather(rate, sum_freq, -setdiff(colnames(.), c("fnr", "fpr", "tnr", "tpr"))) %>%
    dplyr::mutate(trans_or_null = ifelse(rate %in% c("fnr", "tpr"), "Transition", "Null" )) %>%
    mutate(
      rate = recode(
        rate,
        tpr = "% True Positive",
        fpr = "% False Positive",
        fnr = "% False Negative",
        tnr = "% True Negative"
      )
    ) %>%
    # dplyr::filter(metric == "mean_var1") %>%
    ggplot(aes(
      x = crit_value,
      y = sum_freq,
      colour = rate,
      fill = rate
    )) +
    viridis::scale_color_viridis(
      discrete = TRUE,
      option = "magma",
      begin = 0,
      end = 0.9,
      name = ""
    ) +
    viridis::scale_fill_viridis(
      discrete = TRUE,
      option = "magma",
      begin = 0,
      end = 0.9,
      alpha = .5,
      name = ""
    ) +
    geom_bar(position = "stack", stat = "identity") +
    ggh4x::facet_nested( metric ~ regime_switch + trans_or_null,
                         # scales = "free_y",
                         # switch = "y",
                         # axes='all',
                         labeller = label_parsed) +
    labs(y = "Rate", x = latex2exp::TeX("$\\sigma_{crit}$"), title = filename) +
    scale_x_continuous(expand = c(0, 0)
                       # limits = c(min(EWS_warnings_ROC_long$crit_value), max(EWS_warnings_ROC_long$crit_value))
    ) +
    scale_y_continuous(n.breaks = 2, limits = c(0,1))

  return(
    style_plot(pl_ROC) + theme(legend.position = "bottom") +  theme(strip.text.y = element_text(angle = 0))
  )
  # filepath_image = format_path(format_pars(utils::modifyList(
  #   pars_template,
  #   list(
  #     type_output = "figs",
  #     subfolder1 = "ROC",
  #     # subfolder2 = pars$subfolder1,
  #     filename =filename,
  #     file_ext = ".png"
  #   )
  # )))
  # save_plot(style_plot(pl_ROC) + theme(legend.position = "bottom") +  theme(strip.text.y = element_text(angle = 0)),
  #           filepath_image,
  #           w = length(unique(EWS_warnings_ROC_sub$regime_switch)) * 10 + 15,
  #           h = 50,
  #           formats = ".png")
  # print(filepath_image)
}

# pl_ROC = plot_ROC(EWS_warnings_ROC_sub = EWS_warnings_ROC, list())
# print(style_plot(pl_ROC))

pls_ROC = EWS_warnings_ROC%>%
  dplyr::group_by_at("regime_switch")%>%# dplyr::group_keys()
  # dplyr::group_split() %>%
  dplyr::group_map(~ plot_ROC(.x, .y), .keep = TRUE)
print(pls_ROC[[1]])
```


```{r}
pl_AUC = EWS_warnings_AUC %>% 
  dplyr::mutate(AUC = as.numeric(as.character(AUC))) %>%
    # dplyr::filter(metric == "mean_var1") %>%
    ggplot(aes(
      x = sigma_obs_noise,
      y = AUC,
      colour = AUC,
      fill = AUC
    )) +
    viridis::scale_color_viridis(
      # discrete = TRUE,
      option = "magma",
      begin = 0,
      end = 0.9,
      name = ""
    ) +
    viridis::scale_fill_viridis(
      # discrete = TRUE,
      option = "magma",
      begin = 0,
      end = 0.9,
      alpha = .5,
      name = ""
    ) +
    geom_bar(position = "stack", stat = "identity") +
    ggh4x::facet_nested( metric ~ regime_switch,
                         scales = "free",
                         # switch = "y",
                         # axes='all',
                         labeller = label_parsed) #+
    # labs(y = "Rate", x = latex2exp::TeX("$\\sigma_{crit}$"), title = filename) +
    # scale_x_continuous(expand = c(0, 0)
                       # limits = c(min(EWS_warnings_ROC_long$crit_value), max(EWS_warnings_ROC_long$crit_value))
    # ) +
    # scale_y_continuous(n.breaks = 2, limits = c(0,1))
print(style_plot(pl_AUC))
```
## 


```{r}
GLV <- readRDS("C:/Users/kever/Documents/PhD_demos/scripts/firstpkg/detGLV/data/s0.92000-s0.98500-by0.00010/PD_8to16/nr1_T2000_ts0.010_Xsigma0.00000_transition_50transSteps.RDS")
GLV <- readRDS("C:/Users/kever/Documents/PhD_demos/scripts/firstpkg/detGLV/data/s0.92000-s0.98500-by0.00010/PD_2to4/nr1_T2000_ts0.010_Xsigma0.00000_transition_50transSteps.RDS")
GLV <- readRDS("C:/Users/kever/Documents/PhD_demos/scripts/firstpkg/detGLV/data/s0.92000-s0.98500-by0.00010/SUBD_Chaotic_to_Mixed-Periodic1/nr1_T2000_ts0.010_Xsigma0.00000_transition_50transSteps.RDS")
GLV2000 <- readRDS("C:/Users/kever/Documents/PhD_demos/scripts/firstpkg/detGLV/data/s0.92000-s0.98500-by0.00010/SUBD_Chaotic_to_Mixed-Periodic1/nr1_T2000_ts0.010_Xsigma0.00000_transition_25transSteps.RDS")
GLV10000 <- readRDS("C:/Users/kever/Documents/PhD_demos/scripts/firstpkg/detGLV/data/s0.92000-s0.98500-by0.00010/SUBD_Chaotic_to_Mixed-Periodic1/nr1_T10000_ts0.010_Xsigma0.00000_transition_25transSteps.RDS")

df = GLV$df
pl1 = rgl::plot3d(df[,c("X1", "X2", "X3")], 
                  col = df[,c("bifpar_idx")], aspect = TRUE, lwd = 1,
                            clab = c("Sepal", "Width (cm)") # clab is used to change the title of the color legend.

                  # alpha = .01,
                  # main = sprintf("s = %.4f", model_pars[["s"]])
                  )
print(pl1)

```

```{r echo=TRUE}
library(plotly)

fig <- plot_ly(df %>% filter(bifpar_idx %in% c(100, 125, 150, 175)), 
               type="scatter3d", mode = "lines+markers",
               x = ~X1, y = ~X2, z = ~X3,
               # colors = c('#FFE1A1', '#683531', 'pink', 'blue'),
               marker = list(color = ~bifpar_idx, size = 2,
                             colorscale="Rainbow", opacity = .5,
                             # colorscale = c('#FFE1A1', '#683531'),
                             showscale = TRUE))
# fig <- fig %>% 
  # add_markers()
fig <- fig %>% layout(scene = list(xaxis = list(title = 'X1'),
                                   yaxis = list(title = 'X2'),
                                   zaxis = list(title = 'X3'))
                      # annotations = list(
                      #   x = 1.13,
                      #   y = 1.05,
                      #   text = 'Miles/(US) gallon',
                      #   xref = 'paper',
                      #   yref = 'paper',
                      #   showarrow = FALSE
                        # )
)
fig

```











```{r echo=TRUE}
library(plotly)
df = GLV10000$df
m = list(t=50,r=50,b=50,l=50)
scenes = list(xaxis = list(title = 'X1'),
                                   yaxis = list(title = 'X2'),
                                   zaxis = list(title = 'X3'))

fig1 <- plot_ly(df %>% filter(bifpar_idx==50), scene='scene1',
               type="scatter3d", mode = "lines+markers",
               x = ~X1, y = ~X2, z = ~X3,
               marker = list(color = "red", size = 1)) %>% 
  layout(margin = m, 
         scene = scenes)
fig2 <- plot_ly(df %>% filter(bifpar_idx==78), scene='scene2', 
               type="scatter3d", mode = "lines+markers",
               x = ~X1, y = ~X2, z = ~X3,
               marker = list(color = "orange", size = 1)) %>% 
  layout(margin = m, 
         scene = scenes)
fig3 <- plot_ly(df %>% filter(bifpar_idx==103), scene='scene3', 
               type="scatter3d", mode = "lines+markers",
               x = ~X1, y = ~X2, z = ~X3,
               marker = list(color = "blue", size = 1)) %>% 
  layout(margin = m, 
         scene = scenes)
# subplot and define scene
fig <- plotly::subplot(fig1, fig2, fig3, nrows = 1) 
# fig <- fig %>% layout(title = "3D Subplots",
#          scene = list(xaxis = list(title = 'X1'),
#                                    yaxis = list(title = 'X2'),
#                                    zaxis = list(title = 'X3')),
#          scene2 = list(xaxis = list(title = 'X1'),
#                                    yaxis = list(title = 'X2'),
#                                    zaxis = list(title = 'X3')),
#          scene3 =list(xaxis = list(title = 'X1'),
#                                    yaxis = list(title = 'X2'),
#                                    zaxis = list(title = 'X3'))
#         )
fig


fig4 <- plot_ly(df %>% 
                  # filter(bifpar_idx==78, win_nr >= (max(win_nr)*.9)) %>%
                  filter(bifpar_idx==77, win_nr <= (max(win_nr)*.2)) %>%
                  arrange(time_idx), 
               type="scatter3d", mode = "markers",
               x = ~X1, y = ~X2, z = ~X3,
               marker = list(color = "orange", size = 1)
               ) %>% 
  layout(margin = m, 
         scene = scenes)
fig4


```


```{r}

bifpar_idx = 77
X0 = GLV10000$X0s %>% as.data.frame() %>% filter(bifpar_idx == !!bifpar_idx) %>% select(all_of(X_names)) %>% unlist()
s_seq = list(GLV10000$bifpar_list[[bifpar_idx]])
nr_timesteps = 100000
timestep = .01

# Generate timeseries
GLV = bifurcation_ts(
  model = GLV_model,
  model_pars = model_pars,
  bifpar_list = s_seq,
  timestep = timestep,
  X0 = X0,
  nr_timesteps = nr_timesteps,
  downsample_pars = downsample_pars,
  X_names = X_names
)
    
df = GLV$df

fig5 <- plot_ly(df %>% 
                  filter(win_nr >= (max(win_nr)*.96)) %>%
                  # filter(win_nr <= (max(win_nr)*.2)) %>%
                  arrange(time_idx), 
               type="scatter3d", mode = "markers",
               x = ~X1, y = ~X2, z = ~X3, color = ~time_idx,
               marker = list(size = 1,
                             colorscale="Rainbow", opacity = .5,
                             showscale = TRUE)
               ) %>% 
  layout(margin = m, 
         scene = scenes)
fig5  

        
```


These warning signs become more clear when you look in time, in phase space, or in network structure.


** in phase space:

Boundary crisis in time:


** in network structure:
