---
title: "Demo of bifurcationEWS package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Demo of bifurcationEWS package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, echo=FALSE,
  # fig.width=10, fig.height=2, 
  fig.fullwidth=TRUE,
  comment = "#>"
)
rgl::setupKnitr(autoprint = TRUE) # suppress opening new window
```

```{r setup}
# devtools::install_github("KCEvers/bifurcationEWS")
library(bifurcationEWS)
library(ggh4x)
library(dplyr)
library(plotly)
library(ggplot2)
```

### Generating timeseries

```{r demo GLV}
p <- 4
mu <- rep(0, p)
r0 = c(1, .72, 1.53, 1.27)
C0 = matrix(c(1,
              1.09,
              1.52,
              0,
              0,
              1,
              0.44,
              1.36,
              2.33,
              0,
              1,
              0.47,
              1.21,
              0.51,
              0.35,
              1),
            p,
            p,
            byrow = TRUE)
X_names = paste0("X", 1:p)
X0      <- runif(p) %>% setNames(paste0("X", 1:p))
times      <- seq(0, 1000, by = 0.01)
model_pars <- list(r = r0, s = .86, C0 = C0, mu = mu)
df <- deSolve::ode(y = X0, times = times, func = GLV_model, parms = model_pars, method = 'lsoda')
pl_ts = df %>% as.data.frame() %>% ggplot() + geom_point(aes(x = time, y = X1), size = .05)
print(style_plot(pl_ts))
```


What are some more complicated regimes that we can find in the model?

There is no "one" type of chaos - chaotic dynamics can look radically different. 

```{r chaos, eval=FALSE, include=FALSE}
times      <- seq(0, 10000, by = 0.01)
sel_idxs = round(.25*length(times)):length(times)
# Chaos 1
model_pars["s"] = .97
df_chaos1 <- deSolve::ode(y = X0, times = times, func = GLV_model, parms = model_pars)
pl1 = rgl::plot3d(df_chaos1[sel_idxs,c("X1", "X2", "X3")], aspect = TRUE, lwd = 1, alpha = .01, main = sprintf("s = %.4f", model_pars[["s"]]))
print(pl1)

# Chaos 2
model_pars["s"] = .995
df_chaos2 <- deSolve::ode(y = X0, times = times, func = GLV_model, parms = model_pars)
pl2 = rgl::plot3d(df_chaos2[sel_idxs,c("X1", "X2", "X3")], aspect = TRUE, lwd = 1, alpha = .01, main = sprintf("s = %.4f", model_pars[["s"]]))
print(pl2)

# Chaos 3
model_pars["s"] = 1.125
df_chaos3 <- deSolve::ode(y = X0, times = times, func = GLV_model, parms = model_pars)
pl3 = rgl::plot3d(df_chaos3[sel_idxs,c("X1", "X2", "X3")], aspect = TRUE, lwd = 1, alpha = .01, main = sprintf("s = %.4f", model_pars[["s"]]))
print(pl3)

# To display in an R Markdown document:
rglwidget()

# To save to a file:
# htmlwidgets::saveWidget(rgl::rglwidget(width = 520, height = 520), 
#                         file = "3dscatter.html",
#                         libdir = "libs",
#                         selfcontained = FALSE
#                         )
```

### Bifurcation diagram

```{r bifurcation diagram, include=FALSE}
# Generate data with changing bifurcation parameter
downsample_pars = list(type = c("average", "one_sample")[1],
                                                  win_size = 50,
                                                  which_X = c(50, "first", "middle", "last", "random")[1],
                                                  seed = 123)
timestep = .01
nr_timesteps = 500
bifpar_pars = list(
  bifpar_start = 1.4, bifpar_end = .8,
  # bifpar_start = .96, bifpar_end = .97, #bifpar_start = .965, bifpar_end = .985,
                                    pre_steps = 0,
                                  baseline_steps = 0, transition_steps = 251, post_steps = 0)
bifpar_list=do.call(get_bifurcation_range, bifpar_pars)
model_pars = list(r = r0, C0 = C0, mu = mu)

GLV_full = bifurcation_ts(model = GLV_model, model_pars = model_pars,
               bifpar_list = bifpar_list,
               timestep = timestep, nr_timesteps = nr_timesteps,
               downsample_pars = downsample_pars,
               X_names = X_names, 
               # X0 = c(X1=0,X2=0,X3=runif(1), X4=runif(1))
               seed_nr = 100
               )
head(GLV_full$df)
# GLV_full = bifurcationEWS::GLV_full

# Get dataframe with peaks
peaks_df = peaks_bifdiag(GLV_full$df, GLV_full$X_names)
head(peaks_df)
```


```{r plot timeseries, echo=FALSE}
pl_ts = GLV_full$df %>% 
  dplyr::filter(bifpar_idx > 1, bifpar_idx < 3) %>%
  ggplot() + 
  geom_line(aes(x = time_idx, y = X1), linewidth = .1) +
  geom_point(aes(x = time_idx, y = X1), size = .1)
print(style_plot(pl_ts))
```

```{r plot bifurcation diagram, echo=FALSE}
# peaks_df_ = peaks_df %>% dplyr::filter(variable == "X1", bifpar_idx > 1)
pl_peaks = peaks_df %>% ggplot() + geom_point(aes(x = bifpar_idx, y = X), size = .1) +
  labs(x = 'Bifurcation parameter', y = "X1", title = "Transient bifurcation diagram") +
  ggh4x::facet_grid2(variable ~ .)

print(style_plot(pl_peaks))
```

### Finding regime boundaries

How do we find the boundaries between these two regimes? A first approach might be to simply count the number of distinct peaks for each bifurcation parameter. However, this neglects the temporal sequence of these peaks, which is required to be consistenct for a limit cycle. Change point detection algorithms would be another option, yet these are often not suitable for cyclical data.

The code below executes a quite crude yet effective approach. For a number of candidate period lengths $k \in ks$, we cluster the peaks corresponding to the same time in the period. We then compute the within-cluster sum of squares (WCSS) to get a sense of the spread within each cluster. A period length $k$ with minimum WCSS is determined as the best period (we deal with subharmonics in a subsequent step, see code).

Note that although there are much more sophisticated temporal clustering algorithms out there (such as provided in the R package **dtwclust**), we only required a very simple metric to assess how well temporal sequences could be distinguished.

```{r periodicity, echo=TRUE}
regimes_full = find_regimes(GLV_full,
                             thresh_node = .001,
                             thresh_coord_spread = .025,
                             thresh_peak_idx_spread=2,
                             min_length_regime =5,
                             nr_smooth = 0,
                              factor_k = .1)
print(regimes_full$regimes)
print(regimes_full$regime_bounds)
```
 
```{r Plot periods, eval=FALSE, include=FALSE}
# Plot bifurcation diagram coloured according to periodicity
pl_periods = ggplot(regimes_full$period_per_var %>%
                        merge(peaks_df) %>%
           dplyr::filter(variable == "X1")
         ) +
    geom_point(aes(x = bifpar_idx, y = X, col = factor(period)
                   ),
               size = 1.05
               ) +
     ggh4x::facet_grid2(variable ~ .) +
    theme_bw() + viridis::scale_color_viridis(
      discrete=TRUE,
      option = 'turbo',
      begin = 0, end = .8) +
  labs(x = 'Bifurcation parameter', y = "X1", title =  "Transient bifurcation diagram")
print(style_plot(pl_periods))
```


```{r Plot regime switches, eval=FALSE, include=FALSE}
# Plot bifurcation diagram 
pl_regimes = ggplot(peaks_df %>%
           dplyr::filter(variable == "X2")
           # dplyr::filter(min_from == c("coord", "peak_idx")[2])
         ) +
    geom_point(aes(x = bifpar_idx, y = X, 
                   # col = factor(period),
                   ),
               # size = .25
               size = 1.05
               ) +
     ggh4x::facet_grid2(variable ~ .) +
    theme_bw() + viridis::scale_color_viridis(
      discrete=TRUE,
      option = 'turbo',
      begin = 0, end = .8) +
  labs(x = 'Bifurcation parameter', y = "X1", title =  "Transient bifurcation diagram")
print(style_plot(pl_regimes))

pl_regimes +
  geom_vline(data = regimes_full$regime_bounds, aes(
    # xintercept = regime2_start_idx, 
    xintercept = regime1_end_idx,
                                                   col = regime_switch_type))
```



### Generating null and transition models

Now, we might be interested in a select few transitions. To get a sense of the sensitivity and specificity of our ability to detect these transitions, we need both transition and null models. To fairly compare transitions, we make them of the same time length, yet this does also go hand in hand with a change in the rate of the control parameter. 

Say we are interested in the regime switch of *** - how might we simulate this? We already have the range of $s$ for which a particular regime switch occurs. However, the system is also multistable for many parameter ranges, such that a randomly chosen initial condition is not likely to yield the desired regime. To circumvent this issue, we make use of our full GLV timeseries. This timeseries offers both the regime boundaries $s$ and corresponding initial condition needed to simulate a new (much shorter) transition model. In addition, we can now add a proper baseline and transition period. 

```{r include=FALSE}
plot_landscape <- function(df, bifpar_list, bifpar_idx1, bifpar_idx2, bifpar_idx3){
  size_marker = 1
  color1 = 'green'
  color2 = 'orange'
  color3 = 'blue'
  name1 = sprintf("s = %.4f", bifpar_list[[bifpar_idx1]]$s)
  name2 = sprintf("s = %.4f", bifpar_list[[bifpar_idx2]]$s)
  name3 = sprintf("s = %.4f", bifpar_list[[bifpar_idx3]]$s)
  m = list(t=50,r=50,b=50,l=50)
  scenes = list(xaxis = list(title = 'X1'),
                yaxis = list(title = 'X2'),
                zaxis = list(title = 'X3'))

  fig1 <- plot_ly(df %>% filter(bifpar_idx==!!bifpar_idx1), scene = 'scene1', name = name1,
                  type="scatter3d", mode = "markers",
                  x = ~X1, y = ~X2, z = ~X3,
                  marker = list(color = color1, size = size_marker)) %>%
    layout(margin = m,
           scene =  scenes)
  fig2 <- plot_ly(df %>% filter(bifpar_idx==!!bifpar_idx2), scene = 'scene2', name = name2,
                  type="scatter3d", mode = "markers",
                  x = ~X1, y = ~X2, z = ~X3,
                  marker = list(color = color2, size = size_marker)) %>%
    layout(margin = m,
           scene = scenes)
  fig3 <- plot_ly(df %>% filter(bifpar_idx==!!bifpar_idx3), scene = 'scene3',  name = name3,
                  type="scatter3d", mode = "markers",
                  x = ~X1, y = ~X2, z = ~X3,
                  marker = list(color = color3, size = size_marker)) %>%
    layout(margin = m,
           scene = scenes)

  # subplot and define scene
  fig <- plotly::subplot(fig1, fig2, fig3, nrows = 1)
  return(fig)
}

```



```{r echo=FALSE}
# Select regime switch
desired_switch = regimes_full$regime_bounds %>%
  # dplyr::filter(regime_switch_type == "Mixed-Periodic to Chaotic or Transitioning" & regime1 == "Period-16 (X2,X3,X4) AND Period-20 (X1)" & regime2 == "Chaotic or Transitioning")
  dplyr::filter(
    regime1 == "Period-4 (X2,X3,X4) AND Period-6 (X1)" &
      regime2 == "Period-10 (X1) AND Period-8 (X2,X3,X4)"
  )

bifpar_pars_PD_4to8 = list(
  pre_steps = 50,
  baseline_steps = 50,
  transition_steps = 50,
  post_steps = 50,
  bifpar_start = bifpar_list[[desired_switch$regime1_halfway_idx]]$s,
  bifpar_end = bifpar_list[[desired_switch$regime2_start_idx]]$s
)

GLVs = list()
regime_lists = list()

with(bifpar_pars_PD_4to8, {
  for (trans_or_null in c("transition", "null")) {
    if (trans_or_null == "transition") {
      bifpar_pars_PD_4to8_ = bifpar_pars_PD_4to8
    } else if (trans_or_null == "null") {
      bifpar_pars_PD_4to8_ = utils::modifyList(bifpar_pars_PD_4to8, list(bifpar_end = NA))
    }
    X0 = unlist(desired_switch[, GLV_full$X_names])
    
    # Generate data with changing bifurcation parameter
    # GLV = bifurcation_ts(model = GLV_model, model_pars = model_pars,
    #                bifpar_pars = bifpar_pars_PD_4to8,
    #                X0 = X0,
    #                timestep = timestep, nr_timesteps = nr_timesteps,
    #                 downsample_pars = downsample_pars,
    #                X_names = X_names
    #                )
    # Use package-made GLV
    if (trans_or_null == "transition") {
      GLV = GLV_trans
    } else if (trans_or_null == "null") {
      GLV = GLV_null
    }
    head(GLV$df)
    
    pl_ts_trans = GLV$df %>%
      dplyr::filter(bifpar_idx == 2) %>%
      tidyr::gather(variable, value,-setdiff(colnames(.), X_names)) %>%
      ggplot() + geom_point(aes(x = time, y = value), size = .1) +
      ggh4x::facet_grid2(variable ~ .)
    print(style_plot(pl_ts_trans))
    
    # We need to check whether our simulation generated the desired regimes using the same process as above:
    # Get dataframe with peaks
    peaks_df = peaks_bifdiag(GLV$df, GLV$X_names)
    regime_list = find_regimes(
      GLV,
      thresh_node = .001,
      thresh_coord_spread = .025,
      thresh_peak_idx_spread = 2,
      min_length_regime = 5,
      nr_smooth = 0,
      factor_k = .1
    )
    print(regime_list$regimes)
    print(regime_list$regime_bounds)
    
    pl_peaks = peaks_df %>% dplyr::filter(variable == "X1") %>%
      ggplot() + geom_point(aes(x = bifpar_idx, y = X), size = .1) +
      labs(x = 'Bifurcation parameter', y = latex2exp::TeX("$X_1$"), 
           title = "Transient bifurcation diagram") +
      geom_rect(aes(xmin = pre_steps + baseline_steps, 
                    xmax = pre_steps + baseline_steps + transition_steps,
                    ymin = -Inf, ymax = Inf
    ),
    alpha = 0.85, col = "blue", fill = NA,
    linewidth = 1.25) +
            geom_rect(data = regime_list$regime_bounds %>% dplyr::filter(regime1 == "Period-4 (X2,X3,X4) AND Period-6 (X1)"), 
                aes(xmin = regime2_start_idx - 1 - transition_steps, 
                    xmax = regime2_start_idx - 1,
                    ymin = -Inf, ymax = Inf
    ),
    alpha = 0.85, col = "green", fill=NA,
    linewidth = 1.25) 
      # geom_vline(data = data.frame(
      #   xintercept = c(
      #     pre_steps + baseline_steps,
      #     pre_steps + baseline_steps + transition_steps
      #   ),
      #   moment = c("Baseline", "Transition"),
      #   col = rep("Parameter change",2)
      # ),
      # aes(xintercept = xintercept))
    
    print(style_plot(pl_peaks))
     
    # Plot change in phase space
   plot_landscape(
        GLV$df,
        GLV$bifpar_list,
        bifpar_idx1 = pre_steps + baseline_steps,
        bifpar_idx2 = pre_steps + baseline_steps + round(.75 * transition_steps),
        bifpar_idx3 = pre_steps + baseline_steps + transition_steps + 1
      )

   # Store
      GLVs[[trans_or_null]] = GLV
      regime_lists[[trans_or_null]] = regime_list
  }
})
```

The bifurcation diagram above illustrates why it is so difficult to generate regime switches: the exact timing of the regime switch may vary depending on the initial condition and rate of change in the bifurcation parameter. We use the "empirical" (and not the "intended") transition moment for the detection of EWS.

To make the regime switch harder to detect, we add some simple white observational noise.

```{r Adding observational noise}
noise_sigma = .01
GLV_trans$noisy_df = GLV_trans$df %>%
  add_obs_noise(GLV_trans$X_names, noise_mean = 0, noise_sigma = noise_sigma, noise_constant = 0)
pl_noisy = ggplot(data = GLV_trans$df %>% 
  dplyr::filter(bifpar_idx > 5, bifpar_idx < 7) %>%
  tidyr::gather(variable, value, -setdiff(colnames(.), X_names))) +
  geom_point(aes(x = time_idx, y = value), col = 'grey50', alpha = .75, size = .05) +
  geom_point(data = GLV_trans$noisy_df %>% 
                 dplyr::filter(bifpar_idx > 5, bifpar_idx < 7) %>%
               tidyr::gather(variable, value, -setdiff(colnames(.), X_names)), 
  aes(x = time_idx, y = value), col = 'orange3', alpha = .75, size = .05) +
  ggh4x::facet_grid2(variable ~ .) + labs(title = sprintf("sigma = %.2f", noise_sigma))
print(style_plot(pl_noisy))
``` 



## Early Warning Signals

### Computing windowed EWS
```{r Windowed EWS}
uni_metrics = c(
  "mean" = mean,
  "variance" = var,
  "autocorrelation" = get_autocorr,
  "skewness" = moments::skewness,
  "kurtosis" = moments::kurtosis
  # "Smax" = get_Smax
)
multi_metrics = c(
  "meanAbsCrossCorr" = get_conn,
  "largestLambdaCovMatrix" = eigenvalue,
  "spatial_variance" = spatial_variance,
  "spatial_skewness" = spatial_skewness,
  "spatial_kurtosis" = spatial_kurtosis
  # "RQA" = runRQA
)

split_df_EWS = run_bifEWS(GLV_trans$noisy_df, GLV_trans$X_names, uni_metrics, multi_metrics, EWS_args = list())
pl_EWS <- split_df_EWS %>% ggplot() +
    geom_line(aes(x = bifpar_idx, y = value), linewidth = .45, alpha = .5, color ='grey50') +
          geom_point(aes(x = bifpar_idx, y = value), size = .6, color ='grey20') +
  ggh4x::facet_wrap2(. ~ metric, ncol = 3, scales = "free_y",
                     # labeller = labeller(metric = label_func)
                     )
print(style_plot(pl_EWS))


```


### Warnings

At what point do we decide the system is transitioning?

We make use of a simple classical technique that involves **. 
By varying the width of the confidence bands, we get a better understanding of the sensitivity and specificity of the measure. 
The only two settings that are required are the definition of the baseline period and the range of $\sigma_{crit}$ that is tested.

```{r warning}

# Add confidence band
sigma_crit = 2
sigmas_crit = seq(.025, 6, by = 0.025)

# Count baseline back
baseline_idx = round(seq(regime_bounds$baseline_start_idx, regime_bounds$baseline_end_idx))
transition_idx = round(seq(regime_bounds$transition_start_idx, regime_bounds$transition_end_idx))

    # Compute at which critical sigma value a warning occurs
    split_df_EWS_CI = get_warnings(split_df_EWS,
                              baseline_idx,
                              transition_idx,
                                  sigmas_crit = sigmas_crit,
                                  nr_consecutive_warnings = 1)
    
    warning_df = split_df_EWS_CI$warning_df %>%
      rowwise() %>%
      dplyr::mutate(warning_signal = sum(nr_warnings != 0), no_warning_signal = sum(nr_warnings == 0)) %>% ungroup()

    if (pars$trans_or_null == "transition"){
      warning_df = warning_df %>%
        dplyr::rename(nr_tp = warning_signal, nr_fn = no_warning_signal) %>%
        dplyr::mutate(nr_fp = NA, nr_tn = NA)
    } else if (pars$trans_or_null == "null"){
      warning_df = warning_df %>%
        dplyr::rename(nr_tn = no_warning_signal, nr_fp = warning_signal) %>%
      dplyr::mutate(nr_tp = NA, nr_fn = NA)
    }


# Plot
sigma_crit = 2
sz_point_nowarning = .1
sz_point_warning = 1

 pl_warnings <- warnings_trans$winEWS_df %>%
      dplyr::mutate(ymin = mean_w0 - sigma_crit * sd_w0,
                    ymax = mean_w0 + sigma_crit * sd_w0,
                    value_within_bounds = ifelse(value >= ymin & value <= ymax, value, NA),
                    value_outside_bounds_on_time = ifelse((value < ymin | value > ymax) & bifpar_idx > baseline_steps & bifpar_idx <= (baseline_steps + transition_steps), value, NA),
                    value_outside_bounds_baseline = ifelse((value < ymin | value > ymax) & bifpar_idx <= baseline_steps, value, NA),
                    value_outside_bounds_too_late = ifelse((value < ymin | value > ymax) & bifpar_idx >= (baseline_steps + transition_steps), value, NA)
      ) %>%
      ggplot()  +
      geom_ribbon(aes(x = bifpar_idx, ymin = ymin, ymax =ymax), fill="gray80", color = "gray40", alpha = .65) +
     geom_line(aes(x = bifpar_idx, y =  as.numeric(as.character(value))), colour = "gray50", linewidth = .1) +
      geom_point(aes(x = bifpar_idx, y =  as.numeric(as.character(value_within_bounds))), colour = "gray50", size = sz_point_nowarning, alpha = .75) +
      geom_point(aes(x = bifpar_idx, y =  as.numeric(as.character(value_outside_bounds_on_time))), colour = 'red', size = sz_point_warning) +
      geom_point(aes(x = bifpar_idx, y = as.numeric(as.character(value_outside_bounds_baseline))), colour = 'purple', size = sz_point_warning) +
      geom_point(aes(x = bifpar_idx, y = as.numeric(as.character(value_outside_bounds_too_late))), colour = 'blue', size = sz_point_warning) +
      ggh4x::facet_wrap2(metric ~ .,
                         # dir = "v", strip.position = "right",
                         labeller = label_parsed,
                         # strip = ggh4x::strip_nested(bleed = TRUE),
                         ncol = 4,
                         scales = "free_y")
print(pl_warnings)

    
```

### Sensitivity vs. Specificity


```{r AUC}


```


## 

```{r}
GLV <- readRDS("C:/Users/kever/Documents/PhD_demos/scripts/firstpkg/detGLV/data/s0.92000-s0.98500-by0.00010/PD_2to4/nr1_T2000_ts0.010_Xsigma0.00000_transition_25transSteps.RDS")
GLV <- readRDS("C:/Users/kever/Documents/PhD_demos/scripts/firstpkg/detGLV/data/s0.92000-s0.98500-by0.00010/PD_2to4/nr1_T2000_ts0.010_Xsigma0.00000_transition_50transSteps.RDS")
df = GLV$df

split_df = noisy_df %>% as.data.frame() %>% group_by(.data$bifpar_idx) %>% group_split()
df_ = split_df[[14]] 
df_ %>% head

C0 = GLV$model_pars$C0
s = GLV$bifpar_list[[14]]$s
C = C0 * s
diag(C) = diag(C0)
# C %*% state

df1 = df_ %>% arrange(.data$time_idx) %>% select(all_of(X_names)) %>% as.matrix()

stats::cor(df1[,1:2])
stats::cor(df1[1:(nrow(df1)-1),1], df1[2:nrow(df1),2])
stats::cor(df1[1:(nrow(df1)-1),2], df1[2:nrow(df1),1])
library(lmtest)
lmtest::grangertest(X1 ~ X2, order = 1, data = as.data.frame(df1))
lmtest::grangertest(X2 ~ X1, order = 1, data = as.data.frame(df1))

lmtest::grangertest(X2 ~ X3, order = 1, data = as.data.frame(df1))
lmtest::grangertest(X3 ~ X2, order = 1, data = as.data.frame(df1))

full_mi <- function(x){
  # Discretize per column
  x_discr <- infotheo::discretize(x, disc="equalfreq",
                             nbins = 100)

  # Mutual information between all combinations of variables
  D = expand.grid(a = 1:ncol(x), b = 1:ncol(x)) %>% as.data.frame() %>% dplyr::filter(a != b)
  MI_df = plyr::laply(1:nrow(D), function(idx){infotheo::mutinformation(x_discr[,D[idx,"a"]], x_discr[,D[idx,"b"]], method="emp") }) %>% as.matrix() %>% t() %>% magrittr::set_colnames(paste0("MI_", apply(D, 1, paste0, collapse = ""))) %>% as.data.frame()
  MI_df$multi =  infotheo::multiinformation(x_discr,method = "emp")
  MI_df$inter =  infotheo::interinformation(x_discr,method = "emp")
  return(MI_df)
}

full_conn <- function(x){
 stats::cor(x) %>% as.data.frame()%>% unlist %>%as.data.frame(row.names = names(.))%>%t()%>%as.data.frame()
}

lagged_conn <- function(x){
  D = expand.grid(a = 1:ncol(x), b = 1:ncol(x)) %>% as.data.frame() %>% dplyr::filter(a != b)
  conn_df = plyr::laply(1:nrow(D), function(idx){stats::cor(x[1:(nrow(x)-1),D[idx,"a"]], x[2:nrow(x),D[idx,"b"]])}) %>% as.matrix() %>% t() %>% magrittr::set_colnames(paste0("r_", apply(D, 1, paste0, collapse = ""))) %>% as.data.frame()
  
 return(conn_df)
}

var_coeff <- function(x){
  var_mod = vars::VAR(x, p = 1, type = "none")
purrr::map(var_mod$varresult, function(x){as.data.frame(x$coefficients)}) %>%
  do.call(rbind, .) %>% as.data.frame() %>% t() %>% as.data.frame() %>% return()

}
uni_metrics = c("autocorrelation"= get_autocorr)
multi_metrics = c(
  # "full_conn"=full_conn,
                  "VAR" = var_coeff,
                  # "mutual_info" = full_mi,
                  # "lagged_conn"=lagged_conn,
                  "meanAbsCrossCorr" = get_conn,
                  "largestLambdaCovMatrix" = eigenvalue,
                  "spatial_variance" = spatial_variance,
                  "spatial_skewness" = spatial_skewness,
                  "spatial_kurtosis" = spatial_kurtosis)
split_df_EWS = split_df %>%
    lapply(function(df_){run_EWS(df_ %>% arrange(.data$time_idx) %>% select(all_of(X_names)) %>% as.matrix(), uni_metrics, multi_metrics, EWS_args = list())} %>%
             mutate(bifpar_idx = unique(df_$bifpar_idx))) %>%
    do.call(rbind, .) %>% as.data.frame()


# split_df_EWS = run_bifEWS(df, X_names, uni_metrics, multi_metrics, EWS_args = list())
pl_EWS <- split_df_EWS %>%
  tidyr::spread(metric, value) %>%
  dplyr::mutate(VAR_X1.X2.l1_ = VAR_X1.X2.l1 / VAR_X1.X1.l1,
                VAR_X1.X3.l1_ = VAR_X1.X3.l1 / VAR_X1.X1.l1,
                VAR_X1.X4.l1_ = VAR_X1.X4.l1 / VAR_X1.X1.l1
                ) %>% tidyr::gather(metric, value, -c(bifpar_idx)) %>%
  ggplot() +
  geom_point(aes(x=bifpar_idx, y = value), size = .25) +
  geom_line(aes(x=bifpar_idx, y = value), linewidth = .25) +
  ggh4x::facet_wrap2(. ~ metric, ncol = 4, scales = "free_y",
                     # labeller = label_func
                     ) +
  # Baseline
  geom_vline(aes(xintercept = 50 + 50), col = 'purple', linewidth = 1) +
  # Transition
  geom_vline(aes(xintercept = 50 + 50 + 25), col = 'purple', linewidth = 1)
print(style_plot(pl_EWS))


```



## Lyapunov Exponent
```{r}
# Generate data with changing bifurcation parameter

timestep = .01
nr_timesteps = 20000
bifpar_list = get_bifurcation_range(bifpar_start = .85, bifpar_end = .85, pre_steps = 0,
                                  baseline_steps = 0, transition_steps = 1, post_steps = 0)

GLV = bifurcation_ts(model = GLV_model, model_pars = list(r = r0, C0 = C0, mu = mu),
               bifpar_list = bifpar_list,
               timestep = timestep, nr_timesteps = nr_timesteps,
               do_downsample = FALSE,
               # downsample_pars = downsample_pars,
               X_names = X_names)
df = GLV$df
head(df)

# Lyapunov exponent (https://en.wikipedia.org/wiki/Lyapunov_exponent)
df_X = df %>% select(all_of(X_names)) %>% slice(10000:10100) %>% as.matrix()
nrow(df_X)
LAMBDA = (1 / (2* nrow(df_X))) * log(df_X %*% t(df_X) )
eigen(LAMBDA)
```



```{r}
GLV <- readRDS("C:/Users/kever/Documents/PhD_demos/scripts/firstpkg/detGLV/data/s0.92000-s0.98500-by0.00010/PD_8to16/nr1_T2000_ts0.010_Xsigma0.00000_transition_50transSteps.RDS")
GLV <- readRDS("C:/Users/kever/Documents/PhD_demos/scripts/firstpkg/detGLV/data/s0.92000-s0.98500-by0.00010/PD_2to4/nr1_T2000_ts0.010_Xsigma0.00000_transition_50transSteps.RDS")
GLV <- readRDS("C:/Users/kever/Documents/PhD_demos/scripts/firstpkg/detGLV/data/s0.92000-s0.98500-by0.00010/SUBD_Chaotic_to_Mixed-Periodic1/nr1_T2000_ts0.010_Xsigma0.00000_transition_50transSteps.RDS")
GLV2000 <- readRDS("C:/Users/kever/Documents/PhD_demos/scripts/firstpkg/detGLV/data/s0.92000-s0.98500-by0.00010/SUBD_Chaotic_to_Mixed-Periodic1/nr1_T2000_ts0.010_Xsigma0.00000_transition_25transSteps.RDS")
GLV10000 <- readRDS("C:/Users/kever/Documents/PhD_demos/scripts/firstpkg/detGLV/data/s0.92000-s0.98500-by0.00010/SUBD_Chaotic_to_Mixed-Periodic1/nr1_T10000_ts0.010_Xsigma0.00000_transition_25transSteps.RDS")

df = GLV$df
pl1 = rgl::plot3d(df[,c("X1", "X2", "X3")], 
                  col = df[,c("bifpar_idx")], aspect = TRUE, lwd = 1,
                            clab = c("Sepal", "Width (cm)") # clab is used to change the title of the color legend.

                  # alpha = .01,
                  # main = sprintf("s = %.4f", model_pars[["s"]])
                  )
print(pl1)

```

```{r echo=TRUE}
library(plotly)

fig <- plot_ly(df %>% filter(bifpar_idx %in% c(100, 125, 150, 175)), 
               type="scatter3d", mode = "lines+markers",
               x = ~X1, y = ~X2, z = ~X3,
               # colors = c('#FFE1A1', '#683531', 'pink', 'blue'),
               marker = list(color = ~bifpar_idx, size = 2,
                             colorscale="Rainbow", opacity = .5,
                             # colorscale = c('#FFE1A1', '#683531'),
                             showscale = TRUE))
# fig <- fig %>% 
  # add_markers()
fig <- fig %>% layout(scene = list(xaxis = list(title = 'X1'),
                                   yaxis = list(title = 'X2'),
                                   zaxis = list(title = 'X3'))
                      # annotations = list(
                      #   x = 1.13,
                      #   y = 1.05,
                      #   text = 'Miles/(US) gallon',
                      #   xref = 'paper',
                      #   yref = 'paper',
                      #   showarrow = FALSE
                        # )
)
fig

```


```{r echo=TRUE}

fig1 <- plot_ly(df %>% filter(bifpar_idx==100), scene = 'scene1',
               type="scatter3d", mode = "markers",
               x = ~X1, y = ~X2, z = ~X3,
               marker = list(color = "red", size = 1)) %>% 
  layout(margin = list(t=150,r=150,b=150,l=150), 
         scene = list(xaxis = list(title = 'X1'),
                                   yaxis = list(title = 'X2'),
                                   zaxis = list(title = 'X3')))
fig2 <- plot_ly(df %>% filter(bifpar_idx==147), scene = 'scene2',
               type="scatter3d", mode = "markers",
               x = ~X1, y = ~X2, z = ~X3,
               marker = list(color = "orange", size = 1)) %>% 
  layout(margin = list(t=150,r=150,b=150,l=150), 
         scene = list(xaxis = list(title = 'X1'),
                                   yaxis = list(title = 'X2'),
                                   zaxis = list(title = 'X3')))
fig3 <- plot_ly(df %>% filter(bifpar_idx==175), scene = 'scene3',
               type="scatter3d", mode = "markers",
               x = ~X1, y = ~X2, z = ~X3,
               marker = list(color = "blue", size = 1)) %>% 
  layout(margin = list(t=150,r=150,b=150,l=150), 
         scene = list(xaxis = list(title = 'X1'),
                                   yaxis = list(title = 'X2'),
                                   zaxis = list(title = 'X3')))
# subplot and define scene
fig <- plotly::subplot(fig1, fig2, fig3, nrows = 1) 
# fig <- fig %>% layout(title = "3D Subplots",
#          scene = list(xaxis = list(title = 'X1'),
#                                    yaxis = list(title = 'X2'),
#                                    zaxis = list(title = 'X3')),
#          scene2 = list(xaxis = list(title = 'X1'),
#                                    yaxis = list(title = 'X2'),
#                                    zaxis = list(title = 'X3')),
#          scene3 =list(xaxis = list(title = 'X1'),
#                                    yaxis = list(title = 'X2'),
#                                    zaxis = list(title = 'X3'))
#         )
fig

```













```{r echo=TRUE}
library(plotly)
df = GLV10000$df
m = list(t=50,r=50,b=50,l=50)
scenes = list(xaxis = list(title = 'X1'),
                                   yaxis = list(title = 'X2'),
                                   zaxis = list(title = 'X3'))

fig1 <- plot_ly(df %>% filter(bifpar_idx==50), scene='scene1',
               type="scatter3d", mode = "lines+markers",
               x = ~X1, y = ~X2, z = ~X3,
               marker = list(color = "red", size = 1)) %>% 
  layout(margin = m, 
         scene = scenes)
fig2 <- plot_ly(df %>% filter(bifpar_idx==78), scene='scene2', 
               type="scatter3d", mode = "lines+markers",
               x = ~X1, y = ~X2, z = ~X3,
               marker = list(color = "orange", size = 1)) %>% 
  layout(margin = m, 
         scene = scenes)
fig3 <- plot_ly(df %>% filter(bifpar_idx==103), scene='scene3', 
               type="scatter3d", mode = "lines+markers",
               x = ~X1, y = ~X2, z = ~X3,
               marker = list(color = "blue", size = 1)) %>% 
  layout(margin = m, 
         scene = scenes)
# subplot and define scene
fig <- plotly::subplot(fig1, fig2, fig3, nrows = 1) 
# fig <- fig %>% layout(title = "3D Subplots",
#          scene = list(xaxis = list(title = 'X1'),
#                                    yaxis = list(title = 'X2'),
#                                    zaxis = list(title = 'X3')),
#          scene2 = list(xaxis = list(title = 'X1'),
#                                    yaxis = list(title = 'X2'),
#                                    zaxis = list(title = 'X3')),
#          scene3 =list(xaxis = list(title = 'X1'),
#                                    yaxis = list(title = 'X2'),
#                                    zaxis = list(title = 'X3'))
#         )
fig


fig4 <- plot_ly(df %>% 
                  # filter(bifpar_idx==78, win_nr >= (max(win_nr)*.9)) %>%
                  filter(bifpar_idx==77, win_nr <= (max(win_nr)*.2)) %>%
                  arrange(time_idx), 
               type="scatter3d", mode = "markers",
               x = ~X1, y = ~X2, z = ~X3,
               marker = list(color = "orange", size = 1)
               ) %>% 
  layout(margin = m, 
         scene = scenes)
fig4


```


```{r}

bifpar_idx = 77
X0 = GLV10000$X0s %>% as.data.frame() %>% filter(bifpar_idx == !!bifpar_idx) %>% select(all_of(X_names)) %>% unlist()
s_seq = list(GLV10000$bifpar_list[[bifpar_idx]])
nr_timesteps = 100000
timestep = .01

# Generate timeseries
GLV = bifurcation_ts(
  model = GLV_model,
  model_pars = model_pars,
  bifpar_list = s_seq,
  timestep = timestep,
  X0 = X0,
  nr_timesteps = nr_timesteps,
  downsample_pars = downsample_pars,
  X_names = X_names
)
    
df = GLV$df

fig5 <- plot_ly(df %>% 
                  filter(win_nr >= (max(win_nr)*.96)) %>%
                  # filter(win_nr <= (max(win_nr)*.2)) %>%
                  arrange(time_idx), 
               type="scatter3d", mode = "markers",
               x = ~X1, y = ~X2, z = ~X3, color = ~time_idx,
               marker = list(size = 1,
                             colorscale="Rainbow", opacity = .5,
                             showscale = TRUE)
               ) %>% 
  layout(margin = m, 
         scene = scenes)
fig5  

        
```
